\LARGE \bf Machine Translation for Tibetan Buddhist Texts

### 2021-04-15 paper/main.tex
removed tibetan text

\subsection{Model Design}
We used a 6 layer Transformer model that used a multi headed attention mechanism and a switchout transform. Each layer of the transformer model can be divided into an encoder and decoder layer. The encoding layer consists first of a self attention layer, an attention aspect which focuses on the encoded sentence, and then outputs to a feed forward layer. The multi part of the multi headed attention means that each head will encode the information in parallel and then the results will be concatenated and passed to the decoder. Figure one demonstrates this process.
\includegraphics[scale=.4]{Transformer.png}
\caption{\\Figure 1: Overview of a Transformer model}
\\
\par
To prevent overfitting and to augment the data, we added in a switchout transform which will randomly switch out words in the source and target sentence. This also has the added advantage of augmenting the data even further which is especially important for our problem as religious Tibetan has a relatively small corpus, around 30 thousand words. We also attempted to use masking, however we found it to be ineffective. Masking hides tokens from the model by replacing the tokens with a <mask> token. Tibetan tokens are not separated in the same way that latin languages separate their tokens. For example,
\begin{verbatim}
“བྱང་ཆུབ་མངོན་པར་རྫོགས་པར་འཚང་རྒྱ་སྟེ།” 
\end{verbatim}
a line in our test corpus translates to “Having the altruistic intention of excellent bodhicitta.” The Tibetan word is only one token. Therefore, the mask would not help the model learn at all, as the entire sentence would be marked as <mask> and there would not be enough context for our model to discover the meaning. As a result, we did not use any form of masking in our model. Switchout does not create the same problem as their is always a token to take its place. \par
After implementing our transformer model with switchout we tuned hyper parameters to maximize our Transformer model. We manually tuned the following hyperparameters: dropout, label smoothing, rnn size, the number of transformer layers and the number of heads of the attention model. The final configuration used can be found in the appendix. \par
We used four different metrics to evaluate our results: accuracy, predicted score, perplexity and bilingual evaluation understudy (BLEU). The validation accuracy tells us how well we are learning our training data, the predicted score tells us the average log likelihood for each word, the perplexity tells us how well the model predicts the data and the bleu score evaluates the quality of the translation. Together, these metrics provide us with a decent picture of how well our model is performed.

@@

\par
Our six layer transformer model with switch out was unsuccessful at translating religious tibetan to english. The translated output does not match closely enough in either pure accuracy or in the meaning of the passage to be useful in translating religious tibetan texts to english. \par
While superficially, the perplexity appears to demonstrate a strong model, further investigation demonstrates that the low perplexity is a sign of inaccuracy. Perplexity measures the number of possible words that a model could predict for any given sentence. A higher perplexity indicates higher uncertainty. However, the low BLEU score, which measures the quality of the machine translation, has already demonstrated that the accuracy is low. This combined with the perplexity shows the models overconfidence in predicting certain phrases. For example, the tibetan phrase །དེ་ནས་བཅོམ་ལྡན་འདས་ལ་བྱང་ཆུབ་སེམས་དཔའ་སེམས་དཔའ་ཆེན་པོ་བྱམས་པས་འདི་སྐད་ཅེས་གསོལ་ཏོ། translates to “So the bodhisattva mahāsattva Maitreya asked the Blessed One”, but the top five of the predicted translations all are of the form “The Blessed One then said to the [adj] [proper noun]”. Even though this sentence is wrong, the perplexity is low because the model is confident in the first part of the sentence. This phenomenon is repeated throughout the results. Keyphrases from the Tibetan sentence are interpreted wrongly in the same way, and then the model learns to be confident about its predictions which lowers the perplexity. \par
The results demonstrate the flaw in the Transformer model using the current parallel Tibetan to English corpuses. Each token in the Tibetan sentence corresponds to an entire phrase in english rather than a separate word. This complicates the translation, and the model learns parts of the phrases quite well. The validation accuracy is relatively high because the transformer model is good at learning parts of these phrases but it is not good at learning entire sentences or stringing together sentences in context which contributes to the low BLEU score. Overall, the model is capable of learning specific phrases well, but fails at the broad task of creating coherent translations from religious tibetan texts to english.

>>>>

\subsection{Model Design}
We used a 6 layer Transformer model that used a multi headed attention mechanism and a switchout transform. Each layer of the transformer model can be divided into an encoder and decoder layer. The encoding layer consists first of a self attention layer, an attention aspect which focuses on the encoded sentence, and then outputs to a feed forward layer. The multi part of the multi headed attention means that each head will encode the information in parallel and then the results will be concatenated and passed to the decoder. Figure one demonstrates this process.
\includegraphics[scale=.4]{Transformer.png}
\caption{\\Figure 1: Overview of a Transformer model}
\\
\par
To prevent overfitting and to augment the data, we added in a switchout transform which will randomly switch out words in the source and target sentence. This also has the added advantage of augmenting the data even further which is especially important for our problem as religious Tibetan has a relatively small corpus, around 30 thousand words. We also attempted to use masking, however we found it to be ineffective. Masking hides tokens from the model by replacing the tokens with a <mask> token. Tibetan tokens are not separated in the same way that latin languages separate their tokens. For example,
a line in our test corpus translated to “Having the altruistic intention of excellent bodhicitta.” However, the Tibetan word is only one token. Therefore, the mask would not help the model learn at all, as the entire sentence would be marked as <mask> and there would not be enough context for our model to discover the meaning. As a result, we did not use any form of masking in our model. Switchout does not create the same problem as their is always a token to take its place. \par
After implementing our transformer model with switchout we tuned hyper parameters to maximize our Transformer model. We manually tuned the following hyperparameters: dropout, label smoothing, rnn size, the number of transformer layers and the number of heads of the attention model. The final configuration used can be found in the appendix. \par
We used four different metrics to evaluate our results: accuracy, predicted score, perplexity and bilingual evaluation understudy (BLEU). The validation accuracy tells us how well we are learning our training data, the predicted score tells us the average log likelihood for each word, the perplexity tells us how well the model predicts the data and the bleu score evaluates the quality of the translation. Together, these metrics provide us with a decent picture of how well our model is performed.

@@

\par
Our six layer transformer model with switch out was unsuccessful at translating religious tibetan to english. The translated output does not match closely enough in either pure accuracy or in the meaning of the passage to be useful in translating religious tibetan texts to english. \par
While superficially, the perplexity appears to demonstrate a strong model, further investigation demonstrates that the low perplexity is a sign of inaccuracy. Perplexity measures the number of possible words that a model could predict for any given sentence. A higher perplexity indicates higher uncertainty. However, the low BLEU score, which measures the quality of the machine translation, has already demonstrated that the accuracy is low. This combined with the perplexity shows the models overconfidence in predicting certain phrases. For example, one Tibetan phrases expert translation was “So the bodhisattva mahāsattva Maitreya asked the Blessed One”, but the top five of the predicted translations all are of the form “The Blessed One then said to the [adj] [proper noun]”. Even though this sentence is wrong, the perplexity is low because the model is confident in the first part of the sentence. This phenomenon is repeated throughout the results. Keyphrases from the Tibetan sentence are interpreted wrongly in the same way, and then the model learns to be confident about its predictions which lowers the perplexity. \par
The results demonstrate the flaw in the Transformer model using the current parallel Tibetan to English corpuses. Each token in the Tibetan sentence corresponds to an entire phrase in english rather than a separate word. This complicates the translation, and the model learns parts of the phrases quite well. The validation accuracy is relatively high because the transformer model is good at learning parts of these phrases but it is not good at learning entire sentences or stringing together sentences in context which contributes to the low BLEU score. Overall, the model is capable of learning specific phrases well, but fails at the broad task of creating coherent translations from religious tibetan texts to english.
